---
title: "MH3511_Project"
output: 
  html_document:
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---
  
#necessary packages

```{r}
install.packages("gplots")
install.packages('GGally')
install.packages("psych")
library(GGally)
library(gplots)  
library(moments)
library(psych)
```

```{r}
## importing the dataset
data = read.csv("C:\\Users\\thuva\\OneDrive\\Documents\\CODING\\R Codes\\MH3511\\MH3511-Data_Analysis_Project\\dataset\\insurance_data.csv", header = T)
head(data)


```{r}
# Checking if there are any columns with missing values
colSums(is.na(data) | data == "")
```

#From the above results, there is 5 missing values for age and 3 missing values for region and so we decide to remove the rows with missing values

```{r}
# Removing the rows that have missing data in them
data = data[!(rowSums(is.na(data) | data == "") > 0), ]
```

```{r}
# Check that rows with empty values are removed
colSums(is.na(data) | data == "")
```

```{r}
# index column is dropped
data = data[,2:11]
head(data)
```

```{r}
dim(data)
```

## Description and Cleaning of Dataset

```{r}
#printing out plots on new tab
dev.new(width = 12, height = 6)

# summary statistics for the main variable of interest, claim
x1 = seq(0, max(data$claim), 0.1)
y1 = dnorm(x1, mean=mean(data$claim), sd=sd(data$claim))*5000*length(data$claim)
boxplot((data$claim), main = "Boxplot of claim")
hist(data$claim, main = "Histogram of claim", xlab = "claim")
lines(x1,y1, col='red')
print(skewness(data$claim))

```

```{r}
#doing a log transformation for claim, and printing out the different plots
boxplot(log(data$claim), main = "Boxplot of log(claim)")
x2 = seq(min(log(data$claim)), max(log(data$claim))+0.5, 0.1)
y2 = dnorm(x2, mean=mean(log(data$claim)), sd=sd(log(data$claim)))*0.5*length(log(data$claim))
hist(log(data$claim), main = "Histogram of log(claim)", xlab = "claim") # log-transformation
lines(x2, y2, col='blue')
print(skewness(log(data$claim)))

#saving the log transformed claim into data
data$claim <- log(data$claim)
names(data)[names(data) == 'claim'] <- 'log(claim)'

```

```{r}
par(mfrow=c(1,1))
qqnorm(data$`log(claim)`)
qqline(data$`log(claim)`, col='red')

qqnorm(data$`log(claim)`)
qqline(data$`log(claim)`, col='red')
```

```{r}
summary(data$`log(claim)`)
```

```{r}
par(mfrow = c(1,2))
hist(data$age, main = "Age", xlab = "Age")
boxplot(data$age, main = "Age")
```
```{r}
barplot(table(data$age), main = "Number of insured patient by age")
```

```{r}
#analysis of BMI
x = seq(min(data$bmi), max(data$bmi), by = 0.1)
y = dnorm(x, mean = mean(data$bmi), sd = sd(data$bmi)) * length(data$bmi) * 5
hist(data$bmi, main = "BMI", xlab = 'BMI')
lines(x,y, col = 'red')
boxplot(data$bmi, main = "BMI")
```

```{r}

```

```{r}
par(mfrow = c(1,2))
qqnorm(sqrt(data$bmi))
qqline(sqrt(data$bmi), col = 'red')

qqnorm(log(data$bmi))
qqline(log(data$bmi), col = 'red')

shapiro.test((data$bmi))
print(skewness(data$bmi))
```
```{r}
#square root transformation of BMI
par(mfrow = c(1,2))
hist(sqrt(data$bmi), main = "sqrt(BMI)", xlab = "sqrt(BMI)")
x = seq(min(sqrt(data$bmi)), max(sqrt(data$bmi)), by = 0.1)
y = dnorm(x, mean = mean(sqrt(data$bmi)), sd = sd(sqrt(data$bmi))) * length(sqrt(data$bmi)) * 0.2
lines(x,y, col = 'blue')
boxplot(sqrt(data$bmi), main = "sqrt(BMI)")
print(skewness(sqrt(data$bmi)))

##savinig the bmi into data
data$bmi <- sqrt(data$bmi)
names(data)[names(data) == 'bmi'] <- 'sqrt(bmi)'
```
```{r}
model = lm(data$`log(claim)`~ data$`sqrt(bmi)`)
resd = residuals(model)
qqnorm(resd)
qqline(resd, col="red")
```


```{r}
par(mfrow = c(1,1))
x = seq(min(data$bloodpressure), max(data$bloodpressure), by = 0.1)
y = dnorm(x, mean = mean(data$bloodpressure), sd = sd(data$bloodpressure)) * length(data$bloodpressure) * 5
hist(data$bloodpressure, main = "bloodpressure", xlab = "bloodpressure")
lines(x,y, col = 'red')
boxplot(data$bloodpressure, main = "bloodpressure")
outliers_before <- boxplot.stats(data$bloodpressure)$out
length(outliers_before)  # Number of outliers before log transform
print(skewness(data$bloodpressure))

```


```{r}
par(mfrow=c(1,2))
x2 <- seq(min(log(data$bloodpressure)), max(log(data$bloodpressure)), length = 100)
y2 <- dnorm(x2, mean = mean(log(data$bloodpressure)), sd = sd(log(data$bloodpressure)))
hist(log(data$bloodpressure), 
     main = "Histogram of log(data$bloodpressure)", 
     xlab = "log(data$bloodpressure)", 
     col = "gray", 
     freq = FALSE)
lines(x2, y2, col = 'blue', lwd = 2)
boxplot(log(data$bloodpressure))
outliers_after <- boxplot.stats(log(data$bloodpressure))$out
length(outliers_after)  # Number of outliers after log transform
print(skewness(log(data$bloodpressure)))

## saving transformed bp into data
data$bloodpressure <- log(data$bloodpressure)
names(data)[names(data) == 'bloodpressure'] <- 'log(bloodpressure)'
```

```{r}
qqnorm(data$`log(bloodpressure)`)
qqline(data$`log(bloodpressure)`, col = "red")
```

```{r}
model = lm(data$`log(claim)`~data$`log(bloodpressure)`)
resd = residuals(model)
qqnorm(resd)
qqline(resd, col="red")
```

## Statistical Analysis

```{r}
cont_data <- data[, c('log(claim)', 'age', 'sqrt(bmi)', 'log(bloodpressure)')]
mat = cor(cont_data) 
heatmap.2(mat, cellnote=round(mat, 2), trace = "none", notecol="black")


pairs.panels(cont_data,
             method = "pearson",       # Correlation method
             hist.col = "gray",        # Histogram fill color
             density = TRUE, col = 'red',           # Adds red density lines
             ellipses = FALSE,         # Skip correlation ellipses
             scale = FALSE,            # Keeps axis scales consistent
             cex.cor = 1.2)            # Adjusts size of correlation text

```


```{r}
pairs(data[,c(2, 4, 5, 10)])
```

#Relation between claims and gender

```{r}
boxplot(log(data$`log(claim)`)~data$gender, ylab = "log(claim)", xlab = "gender")
```
```{r}
barplot(table(data$gender), main = "Number of insured patients by gender", le)

```
```{r}
var.test(data$`log(claim)`~data$gender)
`````````
```{r}
t.test(data$`log(claim)`[data$gender == "male"],data$`log(claim)`[data$gender == "female"], var.equal = FALSE)

```
#Relation between claim and smoker

```{r}
boxplot(data$`log(claim)`~data$smoker, xlab = "smoker", ylab = "log(claim)")
```
```{r}
barplot(table(data$smoker), main = "Number of insured patients by their smoking status")
```

```{r}
var.test(data$`log(claim)` ~ data$smoker)
```

```{r}
t.test(data$`log(claim)`[data$smoker == "No"], data$`log(claim)`[data$smoker == "Yes"])
```


#Relation between claims and diabetic

```{r}
boxplot(log(data$`log(claim)`)~data$diabetic, xlab = "diabetic", ylab = "log(claim)")
```
```{r}
barplot(table(data$diabetic), main="Number of insured patients by diabetic status")
```

```{r}
var.test(data$`log(claim)` ~ data$diabetic)
```

```{r}
t.test(data$`log(claim)`[data$diabetic == "No"], data$`log(claim)`[data$diabetic == "Yes"], var.equal = TRUE)
```


# Relation between claim and region

```{r}
boxplot(log(data$claim)~data$region, xlab = "region", ylab = "log(claim)")
abline(h=mean(data$logclaim), col='red')
```
```{r}
barplot(table(data$region), main = "Number of insured patient in each region")
```

```{r}
summary(aov(data$`log(claim)` ~ factor(data$region)))
oneway.test(data$`log(claim)` ~ data$region, data = data, var.equal = FALSE)
```
```{r}
pairwise.t.test(data$`log(claim)`, data$region, p.adjust.method="none")
```


# Relation between claim and children

```{r}
boxplot(data$`log(claim)`~data$children, xlab = "children", ylab = "log(claim)")
abline(h = mean(data$`log(claim)`), col = 'blue')
```
```{r}
barplot(table(data$children), main = "Number of insured patients by their number of children")
oneway.test(data$`log(claim)` ~ data$children, data = data, var.equal = F)
pairwise.t.test(data$`log(claim)`, data$children, p.adjust.method = 'none')
sum(data$children == 5)
```

```{r}
summary(aov(data$`log(claim)` ~ factor(data$children)))
```
```{r}
pairwise.t.test(data$`log(claim)`, data$children, p.adjust.method = "none")
```


```{r}
library(ggplot2)

# Plot for sqrt(bmi) to identify threshold
ggplot(data, aes(x = `sqrt(bmi)`, y = `log(claim)`)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 5.45, linetype = "dashed", color = "blue") +
  geom_hline(yintercept = 10.4, linetype = "dashed", color = "red") +
  annotate("text", x = 5.45, y = max(data$`log(claim)`), 
           label = "Threshold = 5.45", hjust = -0.1, color = "blue") +
  annotate("text", x = min(data$`sqrt(bmi)`) + 0.2, y = 10.4, 
           label = "log(claim) = 10.4", vjust = -1, color = "red") +
  labs(title = "log(claim) vs sqrt(bmi)", 
       x = "sqrt(bmi)", 
       y = "log(claim)") +
  theme_minimal()



# Create group based on threshold
data$bmi_group <- ifelse(data$`sqrt(bmi)` > 5.45, "More than 5.45", "Lower than 5.45")
sum(data$bmi_group == "Lower than 5.45")


# Boxplot
boxplot(`log(claim)` ~ bmi_group, data = data,
        main = "log(claim) by sqrt(bmi) group",
        ylab = "log(claim)", xlab = "sqrt(bmi) threshold group")

# t-test to compare means
t.test(`log(claim)` ~ bmi_group, data = data)

# plot for log(bloodpressure) for threshold
ggplot(data, aes(x = `log(bloodpressure)`, y = `log(claim)`)) +
  geom_point(alpha = 0.3) +
  geom_vline(xintercept = 4.61, linetype = "dashed", color = "blue") +
  geom_hline(yintercept = 9.13, linetype = "dashed", color = "red") +
  annotate("text", x = 4.61, y = max(data$`log(claim)`), 
           label = "Threshold = 4.61", hjust = -0.1, color = "blue") +
  annotate("text", x = min(data$`log(bloodpressure)`) + 0.1, y = 10.4, 
           label = "log(claim) = 9.13", vjust = -1, color = "red") +
  labs(title = "log(claim) vs log(bloodpressure)", 
       x = "log(bloodpressure)", 
       y = "log(claim)") +
  theme_minimal()

# Create a new variable based on the 4.61 threshold
data$bp_group <- ifelse(data$`log(bloodpressure)` > 4.61, "More than 4.61", "Less than 4.61")

boxplot(`log(claim)` ~ bp_group, data = data,
        main = "log(claim) by log(bloodpressure) group",
        ylab = "log(claim)", xlab = "log(bloodpressure) threshold group")

```

