---
title: "MH3511_Project"
output: 
  html_document:
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r}
data = read.csv("C:\\Users\\victo\\Downloads\\Insurance_data\\insurance_data.csv", header = T)
head(data)
```

```{r}
colSums(is.na(data) | data == "")
```

From the above results, there is 5 missing values for age and 3 missing values for region and so we decide to remove the rows with missing values

```{r}
data = data[!(rowSums(is.na(data) | data == "") > 0), ]
```

```{r}
# Check that rows with empty values are removed
colSums(is.na(data) | data == "")

```

## Description and Cleaning of Dataset

```{r}
par(mfrow=c(1,3))
# summary statistics for the main variable of interest, claim
boxplot(data$claim)
hist(data$claim)
hist(log(data$claim)) # log-transformation
```

```{r}
par(mfrow = c(1,2))
hist(data$age)
boxplot(data$age)
```

```{r}
par(mfrow = c(1,2))
hist(data$bmi)
boxplot(data$bmi)
```

```{r}
par(mfrow = c(1,2))
hist(data$bloodpressure)
boxplot(data$bloodpressure)
```

## Statistical Analysis

```{r}
cor(data[,c(3, 5, 6, 11)])
```

Relation between claims and gender

```{r}
boxplot(data$claim~data$gender)
```

Relation between claims and diabetic

```{r}
boxplot(data$claim~data$diabetic)
```

Relation between claim and children

```{r}
boxplot(data$claim~data$children)
```

Relation between claim and smoker

```{r}
boxplot(data$claim~data$smoker)
```

Relation between claim and region

```{r}
boxplot(data$claim~data$region)
```
