---
title: "MH3511_Project"
output: 
  html_document:
    fig_caption: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r}
data = read.csv("C:\\Users\\victo\\Downloads\\Insurance_data\\insurance_data.csv", header = T)
head(data)
```

```{r}
colSums(is.na(data) | data == "")
```

From the above results, there is 5 missing values for age and 3 missing values for region and so we decide to remove the rows with missing values

```{r}
data = data[!(rowSums(is.na(data) | data == "") > 0), ]
```

```{r}
# Check that rows with empty values are removed
colSums(is.na(data) | data == "")
```

```{r}
# index column is dropped
data = data[,2:11]
head(data)
```

```{r}
dim(data)
```

## Description and Cleaning of Dataset

```{r}
# summary statistics for the main variable of interest, claim
x1 = seq(0, max(data$claim), 0.1)
y1 = dnorm(x1, mean=mean(data$claim), sd=sd(data$claim))*5000*length(data$claim)
boxplot((data$claim))
hist(data$claim, main = "Histogram of claim", xlab = "claim")
lines(x1,y1, col='red')

```

```{r}
boxplot(log(data$claim), main = "Boxplot of log(claim)")
x2 = seq(min(log(data$claim)), max(log(data$claim))+0.5, 0.1)
y2 = dnorm(x2, mean=mean(log(data$claim)), sd=sd(log(data$claim)))*0.5*length(log(data$claim))
hist(log(data$claim), main = "Histogram of log(claim)", xlab = "claim") # log-transformation
lines(x2, y2, col='blue')

```

```{r}
par(mfrow=c(1,2))
qqnorm(data$claim)
qqline(data$claim, col='red')

qqnorm(log(data$claim))
qqline(log(data$claim), col='red')
```

```{r}
library(moments)
print(skewness(data$claim))
print(skewness(log(data$claim)))
```

```{r}
summary(log(data$claim))
```

```{r}
par(mfrow = c(1,2))
hist(data$age, main = "Age", xlab = "Age")
boxplot(data$age, main = "Age")
```
```{r}
barplot(table(data$age), main = "Number of insured patient by age")
```
```{r}
qqnorm(sqrt(data$age))
qqline(sqrt(data$age))
```

```{r fig4, out.height = 6}
dev.new(width = 12, height = 6)
x = seq(min(data$bmi), max(data$bmi), by = 0.1)
y = dnorm(x, mean = mean(data$bmi), sd = sd(data$bmi)) * length(data$bmi) * 5
hist(data$bmi, main = "BMI", xlab = 'BMI')
lines(x,y, col = 'red')
boxplot(data$bmi, main = "BMI")
```

```{r}

```

```{r}
par(mfrow = c(1,2))
qqnorm(sqrt(data$bmi))
qqline(sqrt(data$bmi), col = 'red')

qqnorm(log(data$bmi))
qqline(log(data$bmi), col = 'red')

shapiro.test((data$bmi))
```
```{r}
par(mfrow = c(1,2))
hist(sqrt(data$bmi), main = "sqrt(BMI)", xlab = "sqrt(BMI)")
x = seq(min(sqrt(data$bmi)), max(sqrt(data$bmi)), by = 0.1)
y = dnorm(x, mean = mean(sqrt(data$bmi)), sd = sd(sqrt(data$bmi))) * length(sqrt(data$bmi)) * 0.2
lines(x,y, col = 'blue')
boxplot(sqrt(data$bmi), main = "sqrt(BMI)")
```
```{r}
model = lm(log(data$claim)~sqrt(data$sqrtbmi))
resd = residuals(model)
qqnorm(resd)
qqline(resd, col="red")
```


```{r}
par(mfrow = c(1,2))
hist(data$bloodpressure, main = "bloodpressure", xlab = "bloodpressure")
boxplot(data$bloodpressure, main = "bloodpressure")
```


```{r}
par(mfrow=c(1,2))
hist(log(data$bloodpressure))
boxplot(log(data$bloodpressure))
shapiro.test((data$bloodpressure))
```

```{r}
qqnorm(sqrt(data$bloodpressure))
qqline(sqrt(data$bloodpressure), col = "red")
```
```{r}
sum(data$bloodpressure > quantile(data$bloodpressure, 0.75) + 1.5*IQR(data$bloodpressure))

sum(data$bloodpressure < quantile(data$bloodpressure, 0.25) - 1.5*IQR(data$bloodpressure))

max(data$bloodpressure)
```


```{r}
model = lm(log(data$claim)~sqrt(data$bloodpressure))
resd = residuals(model)
qqnorm(resd)
qqline(resd, col="red")
```
```{r}
data["logclaim"] = log(data$claim)
data["sqrtbmi"] = sqrt(data$bmi)
```

## Statistical Analysis

```{r}
library(gplots)
mat = cor(data[,c(2, 4, 5, 10)])
heatmap.2(mat, cellnote=round(mat, 2), trace = "none", notecol="black")
```

```{r}
pairs(data[,c(2, 4, 5, 10)])
```

Relation between claims and gender

```{r}
boxplot(log(data$claim)~data$gender, ylab = "log(claim)", xlab = "gender")
```
```{r}
barplot(table(data$gender), main = "Number of insured patients by gender", le)
```
```{r}
var.test(data$logclaim ~ data$gender)
```
```{r}
t.test(data$logclaim[data$gender == "male"], data$logclaim[data$gender =="female"], var.equal = FALSE)
```

Relation between claims and diabetic

```{r}
boxplot(log(data$claim)~data$diabetic, xlab = "diabetic", ylab = "log(claim)")
```
```{r}
barplot(table(data$diabetic), main="Number of insured patients by diabetic status")
```

```{r}
var.test(data$logclaim ~ data$diabetic)
```

```{r}
t.test(data$logclaim[data$diabetic == "No"], data$logclaim[data$diabetic == "Yes"], var.equal = TRUE)
```

Relation between claim and children

```{r}
boxplot(log(data$claim)~data$children, xlab = "children", ylab = "logclaim")
abline(h = mean(data$logclaim), col = 'blue')
```
```{r}
barplot(table(data$children), main = "Number of insured patients by their number of children")
```

```{r}
summary(aov(data$logclaim ~ factor(data$children)))
```
```{r}
pairwise.t.test(data$logclaim, data$children, p.adjust.method = "none")
```


Relation between claim and smoker

```{r}
boxplot(log(data$claim)~data$smoker, xlab = "smoker", ylab = "log(claim)")
```
```{r}
barplot(table(data$smoker), main = "Number of insured patients by their smoking status")
```

```{r}
var.test(data$logclaim ~ data$smoker)
```

```{r}
t.test(data$logclaim[data$smoker == "No"], data$logclaim[data$smoker == "Yes"])
```

Relation between claim and region

```{r}
boxplot(log(data$claim)~data$region, xlab = "region", ylab = "log(claim)")
abline(h=mean(data$logclaim), col='red')
```
```{r}
barplot(table(data$region), main = "Number of insured patient in each region")
```

```{r}
summary(aov(data$logclaim ~ factor(data$region)))
```
```{r}
pairwise.t.test(data$logclaim, data$region, p.adjust.method="none")
```


```{r}
continuous = data[,c(2, 5, 11, 12)]
```

```{r}
library(PerformanceAnalytics)

chart.Correlation(continuous, histogram = TRUE, method = "pearson")
```
```{r}
model = lm(data$logclaim~data$sqrtbmi + log(data$bloodpressure) + data$age + factor(data$smoker) + factor(data$diabetic) + factor(data$region) + factor(data$gender))
step(model, direction = 'backward')
```
```{r}
library(ggplot2)
```

```{r}

```

